---
title: "QMM7300 Case Study 2"
author: "Shairy Bansal A00328001"
date: "2025-08-11"
output: html_document
---

# Introduction

In the fall term in this course i collected personalized data on 7 variables for 89 days.

```{r include=FALSE}
shairy <-read.csv("Shairy-Module 1 - Applied Activity.csv")
nrow(shairy)
```

| Variable     | How is it recorded                   | Type                |
|:-------------|:-------------------------------------|:--------------------|
| Date         | dd/mm/yyyy                           | Identifier          |
| Class        | Yes/No                               | Nominal Category    |
| Study        | Hours                                | Continuous Category |
| Social Media | Yes/No                               | Nominal Category    |
| Energy Level | 1 (very tired) to 5 (very energetic) | Ordinal Category    |
| Steps        | Steps taken                          | Discrete Category   |
| Water Intake | Liters                               | Continuous Category |

The analysis will focus on:

1.  Probability relationships.

2.  Probability distribution.

3.  Confidence intervals.

4.  Hypothesis testing.

# Exploratory Data Analysis

### **1. Contingency Table**

First, I'll create a contingency table for Class attendance and Social Media usage:

```{r echo=FALSE}
# Load the data
data <- read.csv("Shairy-Module 1 - Applied Activity.csv")

# Create contingency table
contingency_table <- table(data$Class, data$Social.Media)
rownames(contingency_table) <- c("No Class", "Class")
colnames(contingency_table) <- c("No Social Media", "Social Media")
contingency_table
```

**2. Probability Analysis**

#### **a. Complement Rule (NOT)**

**Question**: What is the probability of NOT attending class?

```{r echo=FALSE}
# Probability of NOT attending class
p_not_class <- sum(data$Class == "No")/nrow(data)
p_not_class
```

P(Not Class) = `r (p_not_class)` 

Interpretation: There's a 48.3% chance I didn't attend class on any given day during this period.

#### **b. Addition Rule (OR)**

**Question**: What is the probability of attending class OR using social media?

```{r echo=FALSE}
# Probability of Class OR Social Media
p_class <- sum(data$Class == "Yes")/nrow(data)
p_social <- sum(data$Social.Media == "Yes")/nrow(data)
p_both <- sum(data$Class == "Yes" & data$Social.Media == "Yes")/nrow(data)

p_class_or_social <- p_class + p_social - p_both
p_class_or_social
```

P(Class OR Social Media) = P(Class) + P(Social Media) - P(Class AND Social Media) = `r round(p_class_or_social, 3)`

Interpretation: There's a 87.6% chance that on any given day I either attended class or used social media.

#### **c. Conditional Probability (GIVEN)**

**Question**: What is the probability of using social media GIVEN that I attended class?

```{r echo=FALSE}
# Conditional probability
p_social_given_class <- sum(data$Class == "Yes" & data$Social.Media == "Yes")/sum(data$Class == "Yes")
p_social_given_class
```

P(Social Media \| Class) = **`r round(p_social_given_class, 3)`**

Interpretation: On days I attended class, there was a 58.7% chance I also used social media.

#### **d. Multiplication Rule (AND)**

**Question**: What is the probability of attending class AND not using social media?

```{r include=FALSE}
# Probability of Class AND No Social Media
p_class_and_no_social <- sum(data$Class == "Yes" & data$Social.Media == "No")/nrow(data)
p_class_and_no_social
```

P(Class AND No Social Media) = **`r round(p_class_and_no_social, 3)`**

Interpretation: There was a 21.3% chance on any given day that I attended class and didn't use social media.

#### **3. Mutually Exclusive Events**

Two events are mutually exclusive if they cannot occur at the same time (P(A AND B) = 0).

Checking all combinations:

-   P(Class AND Social Media) = 27/89 ≠ 0

-   P(Class AND No Social Media) = 19/89 ≠ 0

-   P(No Class AND Social Media) = 32/89 ≠ 0

-   P(No Class AND No Social Media) = 11/89 ≠ 0

**Conclusion**: There are no mutually exclusive events in this table.

#### **4. Independent Events**

Two events are independent if P(A AND B) = P(A) × P(B).

Let's check if Class attendance and Social Media usage are independent:

```{r echo=FALSE}
# Check independence
p_class <- sum(data$Class == "Yes")/nrow(data)
p_social <- sum(data$Social.Media == "Yes")/nrow(data)
p_class_and_social <- sum(data$Class == "Yes" & data$Social.Media == "Yes")/nrow(data)

is_independent <- abs(p_class_and_social - (p_class * p_social)) < 0.05
is_independent
```

P(Class AND Social Media) = **`r round(p_class_and_social, 3)`**\
P(Class) × P(Social Media) = **`r round(p_class * p_social, 3)`**

Since **`r p_class_and_social`** ≈ **`r p_class * p_social`** (difference = **`r round(abs(p_class_and_social - (p_class * p_social)), 3)`**), we can conclude that class attendance and social media usage are **independent** events.

Interpretation: My social media usage patterns don't appear to be influenced by whether or not I attended class, and vice versa.

### **5. Probability Distribution of Study Hours**

The Canadian University Survey Consortium found that only 12% of students spend more than 4.29 hours per day studying (equivalent to \>30 hours/week).

#### **a. Create Categorical Variable**

```{r echo=FALSE}
# Create categorical variable
data$Study_More_Than_4.29 <- ifelse(data$Study > 4.29, 1, 0)
```

#### **b. Sample Proportion**

```{r echo=FALSE}
# Sample proportion
sample_prop <- mean(data$Study_More_Than_4.29)
sample_prop
```

Sample proportion = **`r round(sample_prop, 3)`** or **`r round(sample_prop * 100, 1)`**%

Interpretation: I studied more than 4.29 hours on **`r round(sample_prop * 100, 1)`**% of days, which is **`r if(sample_prop > 0.12) {"higher"} else {"lower"}`** than the national average of 12%.

#### **6.**

#### **a.Normal Model Parameters**

For a binomial distribution, the mean and standard deviation of the sample proportion are:

-   Mean (μ) = p = 0.12

-   Standard deviation (σ) = sqrt(p(1-p)/n)

```{r echo=FALSE}
# Normal model parameters
n <- nrow(data)
p_pop <- 0.12
mean_prop <- p_pop
sd_prop <- sqrt(p_pop * (1 - p_pop) / n)

c(mean_prop, sd_prop)
```

Mean = `r round(mean_prop, 4)`\
Standard deviation = `r round(sd_prop, 4)`

#### **d. Probability of Sample Proportion**

```{r echo=FALSE}
# Probability of getting sample proportion or less
prob_less <- pnorm(sample_prop, mean = mean_prop, sd = sd_prop)
prob_less
```

P( ≤ **`r round(sample_prop, 3)`**) = **`r round(prob_less, 4)`**

Interpretation: There's a **`r round(prob_less * 100, 1)`**% chance of observing a sample proportion of **`r round(sample_prop * 100, 1)`**% or less if the true population proportion is 12%.

### **7. Confidence Interval for Study Proportion**

```{r echo=FALSE}
# 95% Confidence Interval
z <- qnorm(0.975)
se <- sqrt(sample_prop * (1 - sample_prop) / n)
ci_lower <- sample_prop - z * se
ci_upper <- sample_prop + z * se

c(ci_lower, ci_upper)
```

95% CI = (**`r round(ci_lower, 3)`**, **`r round(ci_upper, 3)`**)

Interpretation: We're 95% confident that the true proportion of days I study more than 4.29 hours is between **`r round(ci_lower * 100, 1)`**% and **`r round(ci_upper * 100, 1)`**%.

### **8. Hypothesis Test**

Test if I study more than the general student population (12% of days \> 4.29 hours).

**Null Hypothesis (H₀)**: p = 0.12\
**Alternative Hypothesis (H₁)**: p \> 0.12 (since sample proportion \> 0.12)

```{r echo=FALSE}
# Hypothesis test
z_score <- (sample_prop - p_pop) / sqrt(p_pop * (1 - p_pop) / n)
p_value <- 1 - pnorm(z_score)

c(z_score, p_value)
```

z-score = **`r round(z_score, 3)`**\
p-value = **`r round(p_value, 4)`**

**Conclusion**: Since the p-value (**`r round(p_value, 4)`**) is **`r if(p_value < 0.05) {"less"} else {"greater"}`** than 0.05, we **`r if(p_value < 0.05) {"reject"} else {"fail to reject"}`** the null hypothesis. This suggests that I **`r if(p_value < 0.05) {"do"} else {"do not"}`** study more than 4.29 hours per day significantly more often than the general Canadian student population.

# Conclusion

Through this analysis, I've learned several things about my class attendance and study habits:

1.  **Class Attendance**: I attended class about 52.8% of the time during this period. My social media usage appears independent of whether I attended class or not.

2.  **Study Habits**: I studied more than 4.29 hours on **`r round(sample_prop * 100, 1)`**% of days, which is **`r if(sample_prop > 0.12) {"higher"} else {"lower"}`** than the national average of 12%. The hypothesis test **`r if(p_value < 0.05) {"confirmed"} else {"did not confirm"}`** that this difference is statistically significant.

**Goals for Next Semester**:

1.  Maintain or improve my study hours while focusing on quality of study time

2.  Monitor social media usage to ensure it doesn't interfere with academic priorities

This analysis has provided valuable insights into my academic habits and will help me make more informed decisions about time management in future semesters.
